---
title: "Ggplot2 and dplyr review"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Main Ideas- Ggplot review

- We had a lot of Docker issues on the ggplot day, so I want to do a review and have you get some more practice.
- We are also going to practice combining ggplot and dplyr.

# Notes

First, load the `tidyverse`,`viridis`, and `fivethirtyeight` packages.

```{r packages, warning = FALSE, message = FALSE}
library(tidyverse)
library(viridis)
library(fivethirtyeight)
```

## Endorsements Data

I am naming an object here so that we can look at the data in the viewer.

```{r load-data, message=FALSE}
endorsements <- endorsements
```

### Step-by-step

Let's practice going step-by-step through a visualization again.

`ggplot` creates the initial base coordinate system that we will add layers to. We first specify the dataset we will use with `data = endorsements`. The `mapping` argument is paired with an aesthetic (`aes`), which tells us how the variables in our dataset should be mapped to the visual properties of the graph.

```{r base-layer}
ggplot(data = endorsements, 
       mapping = aes(x = endorsement_points, y =
                        primary_vote_percentage))
```

```{r add-points}
ggplot(data = endorsements, 
       mapping = aes(x = endorsement_points, y =
                        primary_vote_percentage)) +
   geom_point()
```

```{r add-smooth}
ggplot(data = endorsements, 
       mapping = aes(x = endorsement_points, y =
                        primary_vote_percentage)) +
   geom_point() +
   geom_smooth()
```

Now, let's add some labels.

```{r add-labels}
ggplot(data = endorsements, 
       mapping = aes(x = endorsement_points, y =
                        primary_vote_percentage)) +
   geom_point() +
   geom_smooth() +
   labs(title = "Do Endorsed Candidates Win
        Primaries?", 
        x = "Endorsement Points", y = "Primary Vote 
        Percentage")
```

The procedure used to construct plots can be summarized using the code below.

```{r template, eval = FALSE}
ggplot(data = [dataset], 
       mapping = aes(x = [x-variable], y = [y-variable])) +
   geom_xxx() +
   geom_xxx() + 
  other options
```

## Aesthetics reminder

An aesthetic is a visual property of one of the objects in your plot.

These include:

- shape

- color

- size

- alpha (transparency)

We can map a variable in our dataset to a color, a size, a transparency, and 
so on. The aesthetics that can be used with each `geom_` can be found in the 
documentation.

### Some practice

#### Numeric Variables

Please make a histogram for the number of endorsement points. Remember to include informative labels.

```{r histogram}

```

### Categorical Variables

Please make a bar plot of the number of candidates by party in the dataset and include informative labels.

```{r bar-plot}

```

**Segmented bar plots** can be used to visualize two categorical variables. Let's say you were interested in how many candidates ran for each party by year. 

Now please make a segmented bar plot using `viridis` colors for the number of candidates who ran in each party by year.

```{r segmented-bar-plot}

```

Ask your own research question and make a visualization here:

```{r on-your-own}

```

## Dplyr and ggplot practice problems

1) Please make a bar plot of the number of **Democratic** candidates by year.

```{r combine, warning=FALSE}

```

2) How many candidates raise a lot of money, a medium amount of money, and a small amount of money. Make a segmented barplot showing the number of candidates in each party who raise 1) under 1 million, 2) between 1 and 2 million, and 3) above 3 million. You will need to use a dplyr command first before making the plot.

```{r combine-2, warning=FALSE}

```

**Remember that you put the pipeline operator between the lines for ggplot and a + between lines for dplyr.**

### Additional Resources

### GGplot2
- https://ggplot2.tidyverse.org/
- https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf
- http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
- https://medium.com/bbc-visual-and-data-journalism/how-the-bbc-visual-and-data-journalism-team-works-with-graphics-in-r-ed0b35693535
- https://ggplot2-book.org/
- https://ggplot2.tidyverse.org/reference/geom_histogram.html
- https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
- https://github.com/GraphicsPrinciples/CheatSheet/blob/master/NVSCheatSheet.pdf

## Dplyr

- https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
- https://style.tidyverse.org/